---
title: "Example Report: Introduction to Quarto and R"
format: html
editor: visual
---

## Collaborators

| Name                       | Institution                |
|----------------------------|----------------------------|
| Steven Hiram Rubio Vasquez | Stockholm University + UVG |
| Alex Holmskär              | Stockholm University       |
| Jakub                      | Student council            |
| Saúl Bibiano               | Stockholm University       |
| Ellen1                     |                            |
| Ellen                      |                            |

------------------------------------------------------------------------

## Introduction

Quarto is an open-source scientific publishing system that lets you combine narrative text, code, and output into a single reproducible document. It supports R, Python, and Julia, and can render to HTML, PDF, Word, and more.

This report walks through the basics of writing and running a Quarto document with R.

------------------------------------------------------------------------

## Basic Text Formatting

You can write **bold**, *italic*, and `inline code` directly in the document. Lists work too:

- Item one
- Item two
- Item three

And you can add section headers using `##`, `###`, etc.

------------------------------------------------------------------------

## Running R Code

When you click the **Render** button, all code chunks are executed and their output is embedded in the document.

```{r}
1 + 1
```

You can hide the code and only show the output using `echo: false`:

```{r}
#| echo: false
2 * 2
```

------------------------------------------------------------------------

## A Simple Data Example

R comes with several built-in datasets. Here we load `iris`, which contains measurements of 150 flowers across three species.

```{r}
# Load and preview the data
data(iris)
head(iris)
```

A quick summary of each variable:

```{r}
summary(iris)
```

------------------------------------------------------------------------

## Visualization

We can create a basic plot with base R:

```{r}
plot(
  iris$Sepal.Length, iris$Petal.Length,
  col  = as.integer(iris$Species),
  pch  = 19,
  xlab = "Sepal Length",
  ylab = "Petal Length",
  main = "Sepal vs Petal Length by Species"
)
legend("topleft",
  legend = levels(iris$Species),
  col    = 1:3,
  pch    = 19
)
```

------------------------------------------------------------------------

## Key Takeaways

- Quarto separates content (text) from computation (code chunks).
- `echo: false` lets you hide code while keeping its output.
- Built-in datasets like `iris` are great for quick exploration.
- Rendering produces a self-contained, shareable HTML file.
